### 2023-02-11 doc/manuscript/translocation.qmd
Add text to introduction

-   Important for conservation action, evolutionary rescue not only promotes recoveries of focal populations, but can also provide sources of resistant individuals to reestablish populations extirpated following disease emergence (Mendelson et al. 2019, Joseph and Knapp 2018).

P3: For resistant populations to contribute to landscape-scale species recovery, the fitness benefits gained through selection should manifest beyond the natal environment, allowing for the long-term growth and persistence of incipient populations in novel environments.

@@

-   Something here about previous work with translocations and disease.

-   Simply put, resistant hosts that demonstrate disease resistance, population growth, and long-term viability in non-natal environments provide a means to recover disease-impacted species at a landscape-scale.

P4: Description of MYL frog species complex, impacts of Bd, and evidence of recent recovery. Cite Knapp et al. (2016) to emphasize that frogs in recovering populations are more resistant to Bd than those in Bd-naive populations.

>>>>

-   Evolutionary rescue be difficult to detect, and this complicates our understanding of the potential importance of evolutionary rescue in allowing the persistence and recovery of host populations following disease emergence (Eskew et al. 2021, Lilley et al. 2020). <!--# Maybe move this to the Discussion? -->

-   Important for conservation action, evolutionary rescue not only promotes recoveries of focal populations (Gignous-Wolfsohn et al. 2021), but can also provide sources of resistant individuals to reestablish populations extirpated following disease emergence (Mendelson et al. 2019, Joseph and Knapp 2018).

P3: <!--# In this paragraph, I took a broader approach than what was originally outlined (which focused on fitness benefits extending beyond natal environment). I'm happy to revisit this approach as necessary. -->

<!--# Can we use "reduced susceptibilty" strategically to avoid potential pitfalls associated with resistance, tolerance, and our lack of understanding of what actual mechanism underlies lower Bd loads and higher frog survival?  -->

@@

Given the dramatic impact of disease on global wildlife populations, reintroductions will often be necessary to reestablish extirpated populations at a landscape scale. However, there are several conditions that will determine the feasibility of this much-needed recovery approach. First, populations that survived following disease arrival have evolved increased resistance. Second, increased resistance contributes to higher individual fitness in the presence of disease, and allows positive population growth. Third, the fitness benefits gained through selection for increased resistance manifest beyond the natal environment, allowing for the establishment of populations reintroduced into novel environments. And fourth, reintroduced populations show long-term growth and persistence in a landscape of disease. To date, few of these conditions are documented for any single species impacted by disease, and as a result the reintroduction of resistant hosts as a means to recover species of conservation concern remains largely an untested idea (@scheele2021). Furthermore, a lack of relevant information about disease or a lack of suitable options to overcome the threat of disease has resulted in the failure of numerous reintroduction efforts (@scheele2021, @fellers2007). With the increase in emerging infectious diseases in recent decades (@daszak2000), these failures highlight the critical need to develop effective tools to recover disease-impacted biodiversity worldwide. <!--# Add text mentioning that most discussion of reintroductions/disease has focused on protecting wild populations from diseases vectored by reintroduced individuals? (Viggers et al. 1993, Ewan et al. 2012, Hartley and Sainsbury 2017)  -->

P4: Description of MYL frog species complex, impacts of Bd, and evidence of recent recovery. Cite Knapp et al. (2016) to emphasize that frogs in recovering populations are less susceptible to Bd than those in Bd-naive populations.

The mountain yellow-legged frog, comprised of the sister species *Rana muscosa* and *Rana sierrae*, is emblematic of the unprecedented global declines of amphibians caused by the recently-emerged amphibian chytrid fungus (*Batrachochytrium dendrobatidis*; @scheele2019). Once the most common amphibian in the high elevation portion of the Sierra Nevada mountains (California, USA; @grinnell1924), this frog has disappeared from more than 90% of its historical range during the past century. Due to the severity of its decline and the increasing probability of extinction, both species as now listed as "endangered" under the U.S. Endangered Species Act. Although the introduction of non-native trout into the once-fishless portion of the Sierra Nevada initiated this decline (@bradford1989; @knapp2000), the arrival of *B. dendrobatidis* (Bd) in the 1960s and its subsequent spread (@vredenburg2019) caused additional widespread population extirpations (@vredenburg2010, @rachowicz2006). These Bd-caused declines are fundamentally different than the fish-caused declines. The removal of fish populations is feasible and results in the rapid recovery of frogs. In contrast, Bd appears to persist in habitats even in the absence of amphibian hosts (reference), and therefore represents a permanent alteration of invaded ecosystems that amphibians will need to overcome to reestablish persistent populations.
